\documentclass[../Dissertation.tex]{subfiles}
\begin{document}
\subsubsection{Historical}
\label{section:historical}

Our treatment of this section follows Hull \cite{Hull:2012} Chapter 21, page 474 onwards.

The key feature here is that we don't make probabilistic assumptions (no more Gaussian distributions). 
Instead, we have historical data for our market variables. 
Our view is that what happened in the past is a guide to what will happen in the future.
That is, tomorrow's price change will be sampled from the distribution of price changes in our historical data.

First we must value our portfolio using today's stock prices.
Let us iterate through our historical data and, for each stock, calculate a vector of price changes $\Delta S_i$.
We do this via the same method as shown in equation~\ref{eqn:returns} and once again assume the mean to be zero.
If our data consists of 1001 days of history, then we get 1000 days of price changes.

For each of these price changes $\Delta S_i$, we compute the difference between the current value of our portfolio and the future value our portfolio to build a sample of changes $\Delta\Pi$.
We then sort $\Delta\Pi$ in order of largest to smallest, positive to negative.
If, for example, our confidence level is $c=99\%$.
Our estimate of VaR occurs at the cut-off point in $\Delta\Pi$ at 99\%.
99\% of 1000 samples is 990, so we take $\Delta\Pi_{990}\sqrt{\Delta t}$, where $\Delta t$ is the time horizon.

The algorithm for the Historical method is shown on page~\pageref{alg:historical}.
					\begin{algorithm}
						\caption{Historical method}
						\begin{algorithmic}[1]
						\Procedure{Historical}{}
							\label{alg:historical}							
							\State Value $\Pi^{today}$ from today's $S_i$
							\ForAll {assets $1 \leq i \leq n$}
							\State Calculate vector $\Delta S_i$ from historical data
							\State Apply all $\Delta S_i$ to $S_i$
							\EndFor
							\State Revalue for $\Pi^{tomorrow}$
							\State $\Delta\Pi = \Pi^{tomorrow}-\Pi^{today}$
							\State Sort $\Delta\Pi$ in descending order
							\State VaR $\gets \Delta\Pi_{99\%}\sqrt{\Delta t}$
						\EndProcedure
						\end{algorithmic}
					\end{algorithm}	
\end{document}