\documentclass[../Dissertation.tex]{subfiles}
\begin{document}
\section{How To Use My Project}

\paragraph{Using the .jar}

Our program is \lstinline|VaR.jar|.

Suppose we want to run an experiment where our portfolio consists of a single stock symbol: GOOG at 100 shares and zero options.
We wish to take 5 years of data, with a 1-day time horizon and confidence interval of 95\%.

We execute the \lstinline|.jar| with the command: \lstinline|java -jar VaR.jar GOOG 100 0 5 1 0.95|.
Now, suppose want three stocks in our portfolio, but no options. 
We must use a pipe separated string to separate the stock symbols and stock deltas.
This time our command is 
\begin{lstlisting}[numbers=none]
java -jar VaR.jar GOOG|AAPL|MSFT 100|200|100 0 5 1 0.95
\end{lstlisting}
That is, we assume a portfolio consisting of the symbols GOOG, AAPL and MSFT with 100,200, and 100 shares respectively.
Now we include options on the shares.
The command is now:
\begin{lstlisting}[numbers=none]
java -jar VaR.jar GOOG|AAPL|MSFT 100|200|100 40|30|40 5 1 0.95
\end{lstlisting}
That is we now have 40, 30, 40 put options for GOOG, AAPL, and MSFT.

\paragraph{Compilation}

The \lstinline|.jar| file was compiled using IntelliJ IDEA 2016.3.7 on a machine running Windows 10.
The source files and their dependencies are included in a folder called \lstinline|src| and \lstinline|lib| respectively.

\paragraph{Output}

Running the program will take some time.
On a Core i5, a single experiment can take between 5 and 20 minutes to run, depending on the portfolio.
On \lstinline|linux.cim.rhul.ac.uk|, this will take considerably longer.
Program outputs will be saved to a folder in the directory of the \lstinline|.jar| file.
These will include: 
\lstinline|.csv| files containing the simulated stock prices and put prices from Monte Carlo and Historical simulations;
a \lstinline|.csv| containing the 1000 moments of estimated VaR;
a \lstinline|output.txt| file containing logging;
and lastly \lstinline|rawData.csv| which contains the experimental results.
\end{document}