\documentclass[../Dissertation.tex]{subfiles}
\begin{document}
\section{Professional Issues}

\paragraph{Conflicts of Interest}

With full disclosure, it must be stated that my project supervisor, Dr Kalnishkan, has also been my lecturer in two modules: CS5200 and CS5930. 
As Dr Kalnishkan is the second marker of this dissertation, there presents itself a potential conflict of interest: does prior information regarding my performance in these modules impact the marking of this dissertation in any way?
It does not and cannot. 

In both modules, coursework and exams results were anonymised.
Additionally, it must be taken at face-value that, throughout the length of this project, whether in meetings or in email correspondence, Dr Kalnishkan and I never discussed my results ever.

\paragraph{Inequality}

Economists, such as Thomas Piketty~\cite{Piketty:2014}, have voiced strong concerns regarding the problem of the widening inequality gap between the rich and poor.
As such, one question that must be asked with regards to the technology discussed in this dissertation is: \textit{who does this benefit}?
Perhaps it is more likely to benefit a select few: wealthy individuals who have the means to invest such as or banks, for instance.

Additionally, this technology is depends on access to reliable data.
While we were able to source data for free, there were some restrictions placed on it.
For instance, real-time price data are delayed by 15 minutes in many cases.
But more problematic for us was that we did not have any historical options data.
Ultimately, good financial data is costly and out of reach for most.

On the other hand, let us consider pension funds, for example.
The beneficiaries of pension funds are not just the rich.
Since pension funds typically invest their funds, those managing these type of funds certainly have a fiduciary responsibility to manage risk.

\paragraph{Commercial Use}

In sourcing data from Google Finance, we agree to not use it in a commercial enterprise without prior written consent~\cite{GOOG:Disclaimer}.

\end{document}